Eliminating object prior-bias from sparse-projection tomographic reconstructions

### 2021-10-11 paper/post_tci.tex
Removed the part about RMSE bet 0 to 1 in each case.
Changes in Suppl. material and # views in the main paper

There also exist methods for estimating new changes (in the test)
directly in projection space. In such techniques which can be used in
longitudinal studies, the object-prior is projected in the same set of
views as the current set of test measurements, and a set of projection
differences is computed. These projection differences are essentially
tomographic projections of differences between the unknown test and
the known object-prior. Hence, a variety of reconstruction algorithms
can be used to estimate such a difference image which can then be
added to the object-prior to yield the final estimate of the
test. Such methods have been proposed
in~\cite{Pourmorteza2015,Lee2012}. However, reconstruction of the
difference image will inherently contain sub-sampling artefacts, and
these artefacts will appear in the final reconstruction as there is no
mechanism to mitigate them (unlike the method that we propose). This
can be seen in our supplementary material~\cite{supp_paper},
where we have compared our method with~\cite{Lee2012}.

@@

We observe
that our method reconstructs new structures while simultaneously
reducing streak artefacts. Table~\ref{table:potato_ssim} shows SSIM and RMSE of
the reconstructed new regions using various methods. The ground truth and reconstructions are normalized to lie in [0-1].
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{blue}{blue}{\includegraphics[width=0.22\columnwidth]{../images/potato/post_tci/6_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/tv_zoomed.png}}
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:potato_2D_results}.}
\label{fig:potato_zoomed_2D_results}
\end{figure}
\end{comment}
%-------------------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) of reconstructed potato from various
    methods. 
%The SSIM of ideal reconstruction is 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here. 
The 2D reconstructed images are shown in Fig.~\ref{fig:potato_2D_results}, and 3D reconstructed volumes can be seen in~\cite{supp_paper}.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D, 6 views} & 0.52 (0.36)
& 0.92 (0.06) & \textcolor{red}{0.97 (0.03)} \\ \hline \textbf{3D, 18 views} & 0.58 (0.26) & 0.75 (0.21) & \textcolor{red}{0.84 (0.16)}
\\ \hline
\end{tabular}
\label{table:potato_ssim}
\end{table}

%--------------------------Okra-- results
\subsubsection{\textbf{Okra}}
\label{Sec:okra_spatially_varying}
The test volume was reconstructed from a partial set of $10\%$ of projection views from which
ground truth was reconstructed (refer Sec.~\ref{sec:datasets}). 2D reconstructions of one of the
slices is shown in Fig.~\ref{fig:okra_2D_results}. A complete 3D
reconstruction can be found in in the supplementary material
~\cite{supp_paper}. The red RoI spans 7 consecutive slices where the
test is different from all of the previously scanned objects. %A
%zoomed-in region around the major region of change (red RoI) is shown
%in Fig.~\ref{fig:okra_zoomed_2D_results}.
As in the test, the
reconstruction by spatially-varying method shows the absence of the
deformity and better removal of sub-sampling artefacts when compared
to FDK and TV. This is also seen in the SSIM and RMSE values in
Table~\ref{table:okra_ssim}. The ground truth and reconstructions are normalized to lie in [0-1].

@@

%---------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) within the RoI of reconstructed okra from various
    methods. 
%SSIM of the ideal reconstruction is 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here.  
The 2D reconstructed images are shown in Fig.~\ref{fig:okra_2D_results}, and 3D reconstructed volumes can be seen in~\cite{supp_paper}.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D, 48 views} & 0.46 (0.28)
& 0.50 (0.24) & \textcolor{red}{0.74 (0.15)} \\ 
\hline \textbf{3D, 45 views} & 0.65 (0.22) & 0.70 (0.21) & \textcolor{red}{0.71 (0.19)}
\\ \hline
\end{tabular}
\label{table:okra_ssim}
\end{table}

\subsection{Reconstruction of datasets without imaging model}
\subsubsection{\textbf{Liver:}}
\label{sec:tmh}

Here we show how our technique is useful in a
real-life medical longitudinal study. Our data consists of
successive scans of the liver taken during a radio-frequency ablation
procedure as described in Sec.~\ref{sec:datasets}. Our goal is to to
track the position of the needle in a relatively stationary
background, while simultaneously reducing sub-sampling artefacts.
Specifically, we choose slices 1-6 as our object-prior, and
reconstruct slice 7 from few-views with the specific goal of tracking
the needle and simultaneously reducing
artefacts. Fig.~\ref{fig:tmh_2D_results} 
shows reconstruction of the test
slice from its measurements from only 30 views. The ground truth and reconstructions are normalized to lie in [0-1] and are quantitatively compared using
SSIM and RMSE in Table~\ref{table:tmh_ssim}.

@@

\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) of 2D reconstruction of radio-frequency ablation data from various methods. Values have been computed within the RoI. 
%For ideal reconstruction, SSIM equals 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here.  
The  reconstructed images are shown in Fig.~\ref{fig:tmh_2D_results},}
\begin{tabular}{|l|c|c|c|c|}
\hline \textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline  0.73 (0.04)
& 0.91 (0.09) & \textcolor{red}{0.95 (0.03)} \\ \hline 
\end{tabular}
\label{table:tmh_ssim}
\end{table}
%-----------------------------------------------------------------

@@

The test volume was reconstructed from a partial set of  
 $1.7\%$ of projection views
from which ground truth was reconstructed (refer
Sec.~\ref{sec:datasets}). The selected 3D ground truth of template
volumes, test volume, as well as the 3D reconstructions are shown in
the supplementary material~\cite{supp_paper}. For the sake of
exposition, the red region of interest (RoI) has been culled out from
7 consecutive slices in the 3D volume to indicate new structures;
other changes can be viewed in the video.  2D reconstruction of one of
the slices is shown in Fig.~\ref{fig:sprouts_2D_results}.
Table~\ref{table:sprouts_ssim} shows the improvement in SSIM and RMSE of the
reconstructed new regions as compared to other methods. The ground truth and reconstructions are normalized to lie in [0-1].
%----------------------------------------------------

@@

%----------------------------------------------------
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{yellow}{yellow}{\includegraphics[width=0.22\columnwidth]{../images/sprouts/post_TCI/2D/20_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/tv_zoomed.png}}
%\subcaptionbox{\mbox{Uniform}\\prior}{\includegraphics[width=0.19\columnwidth]{../images/okra/pca_cropped.png}}\hfill
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:sprouts_2D_results}}
\label{fig:sprouts_zoomed_2D_results}
\end{figure}
\end{comment}
%--------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) of reconstructed sprouts from various
    methods. 
%The SSIM of ideal reconstruction) is 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here. 
The 2D reconstructed images are shown in Fig.~\ref{fig:sprouts_2D_results}, and 3D reconstructed volumes can be seen in~\cite{supp_paper}.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} & \textbf{This paper}
 \\ \hline \textbf{2D, 20 views} & 0.67 (0.24)
& 0.82 (0.17) & \textcolor{red}{0.86 (0.13)} \\ \hline \textbf{3D, 30 views} & 0.67 (0.25) & 0.73 (0.23) & \textcolor{red}{0.82 (0.14)}
\\ \hline
\end{tabular}
\label{table:sprouts_ssim}
\end{table}

>>>>

There also exist methods for estimating new changes (in the test)
directly in projection space. In such techniques which can be used in
longitudinal studies, the object-prior is projected in the same set of
views as the current set of test measurements, and a set of projection
differences is computed. These projection differences are essentially
tomographic projections of differences between the unknown test and
the known object-prior. Hence, a variety of reconstruction algorithms
can be used to estimate such a difference image which can then be
added to the object-prior to yield the final estimate of the
test. Such methods have been proposed
in~\cite{Pourmorteza2015,Lee2012}. However, reconstruction of the
difference image will inherently contain sub-sampling artefacts, and
these artefacts will appear in the final reconstruction as there is no
mechanism to mitigate them (unlike the method that we propose). This
can be seen in Sec.~3 our supplementary material~\cite{supp_paper},
where we have compared our method with~\cite{Lee2012}.

@@

We observe
that our method reconstructs new structures while simultaneously
reducing streak artefacts. Table~\ref{table:potato_ssim} shows SSIM and RMSE of
the reconstructed new regions using various methods. 
%The ground truth and reconstructions are normalized to lie in [0-1].
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{blue}{blue}{\includegraphics[width=0.22\columnwidth]{../images/potato/post_tci/6_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/tv_zoomed.png}}
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:potato_2D_results}.}
\label{fig:potato_zoomed_2D_results}
\end{figure}
\end{comment}
%-------------------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) within the RoI of reconstructed potato from various
    methods. 
%The SSIM of ideal reconstruction is 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here. 
The 2D reconstructed images are shown in Fig.~\ref{fig:potato_2D_results}, and 3D reconstructed volumes can be seen in~\cite{supp_paper}.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D} & 0.52 (0.36)
& 0.92 (0.06) & \textcolor{red}{0.97 (0.03)} \\ \hline \textbf{3D} & 0.58 (0.26) & 0.75 (0.21) & \textcolor{red}{0.84 (0.16)}
\\ \hline
\end{tabular}
\label{table:potato_ssim}
\end{table}

%--------------------------Okra-- results
\subsubsection{\textbf{Okra}}
\label{Sec:okra_spatially_varying}
The test volume was reconstructed from a partial set of $10\%$ of projection views from which
ground truth was reconstructed (refer Sec.~\ref{sec:datasets}). 2D reconstructions of one of the
slices is shown in Fig.~\ref{fig:okra_2D_results}. A complete 3D
reconstruction can be found in in the supplementary material
~\cite{supp_paper}. The red RoI spans 7 consecutive slices where the
test is different from all of the previously scanned objects. %A
%zoomed-in region around the major region of change (red RoI) is shown
%in Fig.~\ref{fig:okra_zoomed_2D_results}.
As in the test, the
reconstruction by spatially-varying method shows the absence of the
deformity and better removal of sub-sampling artefacts when compared
to FDK and TV. This is also seen in the SSIM and RMSE values in
Table~\ref{table:okra_ssim}. 
%The ground truth and reconstructions are normalized to lie in [0-1].

@@

%---------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) within the RoI of reconstructed okra from various
    methods. 
%SSIM of the ideal reconstruction is 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here.  
The 2D reconstructed images are shown in Fig.~\ref{fig:okra_2D_results}, and 3D reconstructed volumes can be seen in~\cite{supp_paper}.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D} & 0.46 (0.28)
& 0.50 (0.24) & \textcolor{red}{0.74 (0.15)} \\ 
\hline \textbf{3D} & 0.65 (0.22) & 0.70 (0.21) & \textcolor{red}{0.71 (0.19)}
\\ \hline
\end{tabular}
\label{table:okra_ssim}
\end{table}

\subsection{Reconstruction of datasets without imaging model}
\subsubsection{\textbf{Liver}}
\label{sec:tmh}

Here we show how our technique is useful in a
real-life medical longitudinal study. Our data consists of
successive scans of the liver taken during a radio-frequency ablation
procedure as described in Sec.~\ref{sec:datasets}. Our goal is to to
track the position of the needle in a relatively stationary
background, while simultaneously reducing sub-sampling artefacts.
Specifically, we choose slices 1-6 as our object-prior, and
reconstruct slice 7 from few-views with the specific goal of tracking
the needle and simultaneously reducing
artefacts. Fig.~\ref{fig:tmh_2D_results} 
shows reconstruction of the test
slice from its measurements from only 30 views. 
The ground truth and reconstructions 
%are normalized to lie in [0-1] and 
are quantitatively compared using
SSIM and RMSE in Table~\ref{table:tmh_ssim}.

@@

\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) within the RoI of 2D reconstruction of radio-frequency ablation data of liver from various methods.  
%For ideal reconstruction, SSIM equals 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here.  
The  reconstructed images are shown in Fig.~\ref{fig:tmh_2D_results},}
\begin{tabular}{|l|c|c|c|c|}
\hline \textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline  0.73 (0.04)
& 0.91 (0.09) & \textcolor{red}{0.95 (0.03)} \\ \hline 
\end{tabular}
\label{table:tmh_ssim}
\end{table}
%-----------------------------------------------------------------

@@

The test volume was reconstructed from a partial set of  
 $1.7\%$ of projection views
from which ground truth was reconstructed (refer
Sec.~\ref{sec:datasets}). The selected 3D ground truth of template
volumes, test volume, as well as the 3D reconstructions are shown in
the supplementary material~\cite{supp_paper}. For the sake of
exposition, the red region of interest (RoI) has been culled out from
7 consecutive slices in the 3D volume to indicate new structures;
other changes can be viewed in the video.  2D reconstruction of one of
the slices is shown in Fig.~\ref{fig:sprouts_2D_results}.
Table~\ref{table:sprouts_ssim} shows the improvement in SSIM and RMSE of the
reconstructed new regions as compared to other methods. 
%The ground truth and reconstructions are normalized to lie in [0-1].
%----------------------------------------------------

@@

%----------------------------------------------------
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{yellow}{yellow}{\includegraphics[width=0.22\columnwidth]{../images/sprouts/post_TCI/2D/20_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/tv_zoomed.png}}
%\subcaptionbox{\mbox{Uniform}\\prior}{\includegraphics[width=0.19\columnwidth]{../images/okra/pca_cropped.png}}\hfill
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:sprouts_2D_results}}
\label{fig:sprouts_zoomed_2D_results}
\end{figure}
\end{comment}
%--------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM (RMSE) within the RoI of reconstructed sprouts from various
    methods. 
%The SSIM of ideal reconstruction) is 1. The ideal RMSE is 0, and the worst RMSE possible is 1 here. 
The 2D reconstructed images are shown in Fig.~\ref{fig:sprouts_2D_results}, and 3D reconstructed volumes can be seen in~\cite{supp_paper}.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} & \textbf{This paper}
 \\ \hline \textbf{2D} & 0.67 (0.24)
& 0.82 (0.17) & \textcolor{red}{0.86 (0.13)} \\ \hline \textbf{3D} & 0.67 (0.25) & 0.73 (0.23) & \textcolor{red}{0.82 (0.14)}
\\ \hline
\end{tabular}
\label{table:sprouts_ssim}
\end{table}
